Rarefied Null Models

### 2021-09-17 analysis/03_rarefy_null_models.Rmd
clearer annotation

Get dataframe of filenames and corresponding rarefyID's
```{r}
path <- here::here("data", "rarefied_samples")
files <- list.files(path) %>%
  paste0(path, "/", .)
```

@@

species_pools <- bt_traitfiltered %>%
  select(rarefyid, species) %>% 
  distinct()

@@

Simulation functions
```{r}
get_plot_sample <- function(file, n){
  #print(filter_id)
  load(file)
  comm <- rare_comm_save %>%
    pivot_longer(cols = starts_with("ITIS"), names_to = "species", values_to = "weight") %>%
    filter(weight != 0) %>%
    rename_with(tolower)

  filter_rarefyid <- unique(comm$rarefyid) %>% unlist()
  samp_rarefy_resamp <- unique(comm$rarefy_resamp) %>% unlist()
  print(filter_rarefyid)

@@

  # plot_data <- data %>% filter(rarefyID == filter_id) %>%
  #   select(year, id, weight)

  species_pool <- species_pools %>%
    filter(rarefyid == filter_rarefyid) %>%
    pull(species) %>%
    na.omit() %>%
    unique() #get column of unique scientific names

  get_samp_species <- function(data){
    samp_species <- sample(species_pool, n_distinct(data$species))
    samp_abun <- sample(data$weight)

@@

  site_samples <- data.frame()
  for (j in 1:n){
    iter_samples <- data.frame()
    for(i in unique(comm$year)){
      year_data <- comm %>% filter(year == i)
      iter_samples <- rbind(iter_samples, get_samp_species(year_data))
    }
    site_samples <- rbind(site_samples, iter_samples %>% mutate(n = j))
  }
  plot_samples <- site_samples %>%
    mutate(rarefyid = filter_rarefyid, rarefy_resamp = samp_rarefy_resamp) %>%
    pivot_wider(names_from = samp_species, values_from = samp_abun)

  save_dir <- here::here("data", "null_samples")
  fs::dir_create(save_dir)
  write.csv(plot_samples, paste0(save_dir, "/", filter_rarefyid, "_samp", samp_rarefy_resamp, ".csv"))

@@

Get sample dataframe
```{r}
future::plan(multiprocess, workers = 30)
sample_occurrence <- future_map(files, get_plot_sample, n = 500, 
                                .options = furrr_options(seed = TRUE)) #%>%
  # bind_rows() %>%
  # replace(is.na(.), 0)
```

@@

```{r}
get_nullplot_FD <- function(file_path, traits, ...) {
  library(FD)

@@

  sample_species_mat <- plot_samples %>%
    select(starts_with("ITIS")) %>% #want only species x site info
    select(sort(tidyselect::peek_vars())) #sort columns to be in alphabetical order, required for dbFDI()

  #print(head(sample_species_mat))
  trait_mat <- get_traitMat(colnames(sample_species_mat), trait_data = traits)

@@

  max_val <- max(unlist(map(sample_species_mat, max)))
  if(max_val > 1){
    FD_mets <- get_FD_safe(species_mat = sample_species_mat, trait_mat = trait_mat, year_list = plot_samples$year,
                           data_id = unique(plot_samples$rarefyid), samp_id = unique(plot_samples$rarefy_resamp),
                           w.abun = TRUE, m = "min", corr = "cailliez")
  }else{
    FD_mets <- get_FD_safe(species_mat = sample_species_mat, trait_mat = traits, year_list = plot_samples$year,
                           data_id = unique(plot_samples$rarefyid), samp_id = unique(plot_samples$rarefy_resamp),
                           w.abun = FALSE, m = "min", corr = "cailliez")
  }

@@

null_files <-  dir(here::here("data", "null_samples"), "*.csv") %>%
  paste0(here::here("data", "null_samples"), "/", .)

@@

Create Null table
```{r}
path <- here::here("data", "fd_null_output")
null_files <-  dir(path, "*.tsv") %>%
  paste0(path, "/", .)

process_nulls <- function(file_path, n){
  data <- read_tsv(file_path) %>%
    rename(richness = nbsp) %>%
    select(-sing.sp, -n) %>%
    group_by(year)

@@

usethis::use_data(null_table)
pins::pin(null_table, board = "github")
```

>>>>

Get dataframe of filenames and corresponding rarefyID's
```{r}
#get list of files
path <- here::here("data", "rarefied_samples")
files <- list.files(path) %>%
  paste0(path, "/", .)
```

@@

#for each rarefyID, get complete list of species
species_pools <- bt_traitfiltered %>%
  select(rarefyid, species) %>% 
  distinct()

@@

Simulation functions
```{r}

#for each file, sample n times; each file is data for one community
get_plot_sample <- function(file, n){
  #print(filter_id)
  load(file) #loads file called rare_comm_save
  #put in matrix form
  comm <- rare_comm_save %>%
    pivot_longer(cols = starts_with("ITIS"), names_to = "species", values_to = "weight") %>%
    filter(weight != 0) %>%
    rename_with(tolower)

  #get rarefyid and resample number 
  filter_rarefyid <- unique(comm$rarefyid) %>% unlist()
  samp_rarefy_resamp <- unique(comm$rarefy_resamp) %>% unlist()
  print(filter_rarefyid)

@@

  #only species occuring at that rarefyid should be in the species pool
  species_pool <- species_pools %>%
    filter(rarefyid == filter_rarefyid) %>%
    pull(species) %>%
    na.omit() %>%
    unique() #get column of unique scientific names

  #function to get sample species and abundances
  get_samp_species <- function(data){
    samp_species <- sample(species_pool, n_distinct(data$species))
    samp_abun <- sample(data$weight)

@@

  #loop over the number of samples
  site_samples <- data.frame()
  for (j in 1:n){
    iter_samples <- data.frame()
    #for each year get a sample species list
    for(i in unique(comm$year)){
      year_data <- comm %>% filter(year == i)
      iter_samples <- rbind(iter_samples, get_samp_species(year_data))
    }
    site_samples <- rbind(site_samples, iter_samples %>% mutate(n = j))
  }
  plot_samples <- site_samples %>%
    mutate(rarefyid = filter_rarefyid, rarefy_resamp = samp_rarefy_resamp) %>%
    pivot_wider(names_from = samp_species, values_from = samp_abun)

  #save out samples
  save_dir <- here::here("data", "null_samples")
  fs::dir_create(save_dir)
  write.csv(plot_samples, paste0(save_dir, "/", filter_rarefyid, "_samp", samp_rarefy_resamp, ".csv"))

@@

Get sample dataframe
```{r}
#get null samples for each rarefied sample
future::plan(multiprocess, workers = 30)
sample_occurrence <- future_map(files, get_plot_sample, n = 500, 
                                .options = furrr_options(seed = TRUE)) 
```

@@

Calculated FD metrics for all the null samples
```{r}
get_nullplot_FD <- function(file_path, traits, ...) {
  library(FD)

@@

  #get species matrix
  sample_species_mat <- plot_samples %>%
    select(starts_with("ITIS")) %>% #want only species x site info
    select(sort(tidyselect::peek_vars())) #sort columns to be in alphabetical order, required for dbFDI()

  #get trait matrix
  trait_mat <- get_traitMat(colnames(sample_species_mat), trait_data = traits)

@@

  # check if data are counts or abundances, then calculated FD
  max_val <- max(unlist(map(sample_species_mat, max)))
  if(max_val > 1){
    FD_mets <- get_FD_safe(species_mat = sample_species_mat, trait_mat = trait_mat, year_list = plot_samples$year,
                           data_id = unique(plot_samples$rarefyid), samp_id = unique(plot_samples$rarefy_resamp),
                           w.abun = TRUE, m = "min", corr = "cailliez")
  }else{
    FD_mets <- get_FD_safe(species_mat = sample_species_mat, trait_mat = traits, year_list = plot_samples$year,
                           data_id = unique(plot_samples$rarefyid), samp_id = unique(plot_samples$rarefy_resamp),
                           w.abun = FALSE, m = "min", corr = "cailliez")
  }

@@

#get list of null sample files
null_files <-  dir(here::here("data", "null_samples"), "*.csv") %>%
  paste0(here::here("data", "null_samples"), "/", .)

@@

Create Null table
```{r}
#get list of null fd sample files
path <- here::here("data", "fd_null_output")
null_files <-  dir(path, "*.tsv") %>%
  paste0(path, "/", .)

#function to read in null fd samples and get info to calculate stats
process_nulls <- function(file_path, n){
  data <- read_tsv(file_path) %>%
    rename(richness = nbsp) %>%
    select(-sing.sp, -n) %>%
    group_by(year)

@@

#write out null table
usethis::use_data(null_table)
pins::pin(null_table, board = "github")
```
